---
title: "Chemotherapy during pregnancy"
output: 
  html_document:
    number_sections: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  include = TRUE, warning = FALSE, message = FALSE, echo = FALSE
)
```

```{r import data, include = FALSE}
packages_list <- c(
  "data.table", "tidyverse", "kableExtra", "PSW", 
  "MASS", "splines", "mice"
) # 'MatchIt',
new_packages <- packages_list[
  !(packages_list %in% installed.packages()[, "Package"])
]
if (length(new_packages)) install.packages(new_packages)
suppressMessages(lapply(packages_list, library, character.only = TRUE))

select <- dplyr::select
rbind <- base::rbind
kable <- knitr::kable

project_folder <- file.path(
  "C:/Users/sren/Dropbox (Partners HealthCare)/BOC shared",
  "Chemo during pregnancy (Sella)"
)

# load functions needed
source(file.path(project_folder, "r", "functions.R"))

data_path <- file.path(project_folder, "data", "2020-12-9")
load(file.path(data_path, "cleaned_data.RData"))

ps_data <- redcap_data[
  chemopg == 1 & pregnancy_outcome == "single", 
]
```

# Propensity score matching 
<!-- {.tabset .tabset-fade .tabset-pills} -->
There are two exposures we want to build models for: taxol and GCSF. Only patients who received chemotherapy and gave single birth were included. Possible covariates include: age at diagnosis (`agedx`), gestational age at diagnosis (`gadx`), tumor stage (`I_stage_34`), breast cancer type (`I_TNBC`, `I_HER2`), timing of chemotherapy (`I_neoadjuvant`) and surgery during pregnancy (`I_surgery`). According to Tal, there are several variables that are related to exposure/outcome:

* GCSF - AC chemotherapy
* Taxol - gestational age at diagnosis, tumor stage
* Outcomes - gestational age

## Propensity score model for Taxol

For patients who received chemotherapy, there is only one patient received taxol without AC. Adding AC as a covariATE makes the standardized difference very large.
```{r taxol and acpreg}
ps_data[, .N, by = c("taxolpreg", "acpreg")] %>% 
  knitr::kable(align = "l", digits = 2, booktabs = TRUE) %>% 
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"), 
    full_width = F, 
    position = "center"
  )
```

```{r modify table 1 vars}
# variables in Table 1:
# age, gadx, yeardx, trimester, cancer stage, cancer type, BRCA1/2 status (mutation), timing (neo/adjuvant), surgery

# collapse factor levels, normalized numeric vars, and splined them
# normalized age
ps_data[, age_norm := (agedx - mean(agedx)) / sd(agedx)]
# spline age
age_norm_summary <- summary(ps_data$age_norm)
age_norm_spline <- bs(ps_data$age_norm, knots = age_norm_summary[c(2, 3, 5)]) # 1st quantile, median and 3rd quantile
colnames(age_norm_spline) <- paste0("age_norm_bs", 1:6)
ps_data <- cbind(ps_data, age_norm_spline)
# normalized gestational age
ps_data[, gadx_norm := (gadx - mean(gadx)) / sd(gadx)]
# spline gestational age
gadx_norm_summary <- summary(ps_data$gadx_norm)
gadx_norm_spline <- bs(ps_data$gadx_norm, knots = gadx_norm_summary[c(2, 3, 5)])
colnames(gadx_norm_spline) <- paste0("gadx_norm_bs", 1:6)
ps_data <- cbind(ps_data, gadx_norm_spline)
# centered year of diagnosis
ps_data[, yeardx_centered :=
  (yeardx - min(yeardx)) / (max(yeardx) - min(yeardx))]
# cancer stage
ps_data[, I_stage_34 := ifelse(stage %in% 3:4, 1, 0)] # collapse 1-2, and 3-4
# cancer type
ps_data[, I_TNBC := ifelse(cancer_type == "TNBC", 1, 0)]
ps_data[, I_HER2 := ifelse(cancer_type == "HER2", 1, 0)]
# BRCA1/2 
ps_data[, I_mutation := ifelse(mutation == "Y", 1, 0)] # collapse 0 and unknown
# chemotherapy timing
ps_data[, I_neoadjuvant := ifelse(chemo_timing == "neoadjuvant", 1, 0)]
# surgery
ps_data[, I_surgery := surgpg]
#  "yeardx_centered",
tbl1_vars <- c(
  "age_norm", "gadx_norm", "yeardx_centered", "I_stage_34", 
  "I_TNBC", "I_HER2", "I_neoadjuvant", "I_surgery"
) # used for std.diff check, not covariates

tbl1_splined_vars <- c(
  paste0("age_norm_bs", 1:6), paste0("gadx_norm_bs", 1:6), 
  "I_stage_34", "I_TNBC", "I_HER2", "I_mutation", 
  "I_neoadjuvant", "I_surgery"
)
```

When only including the following covariates in the propensity score model, the standardized differences for other variables in Table 1 do not look very good:
\[
logit(taxol) \sim I(stage = 3, 4) * gestational~age \\
logit(taxol) \sim I(stage = 3, 4) + splined(gestational~age)
\]

```{r w/o spline few covariates , out.width = '70%', fig.show = 'hide', fig.align = 'center', eval = FALSE}
# # Propensity score matching using ATE weight
# psw(
#   data = as.data.frame(ps_data),
#   form.ps = taxolpreg ~ I_stage_34 * gadx_norm,
#   weight = 'ATE',
#   std.diff = TRUE,
#   V.name = tbl1_vars
# ) %>%
#   extractStdDiff(.) %>% 
#   knitr::kable(align = "l", digits = 2, booktabs = TRUE,
#                caption = "ATE weighted standardized difference") %>% 
#   kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F, position = "center")

# Propensity score matching using MW weight
psw(
  data = as.data.frame(ps_data),
  form.ps = taxolpreg ~ I_stage_34 * gadx_norm,
  weight = "MW",
  std.diff = TRUE,
  V.name = tbl1_vars
) %>%
  extractStdDiff(.) %>%
  knitr::kable(
    align = "l", digits = 2, booktabs = TRUE,
    caption = "MW weighted std.diff with few covariATEs"
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"), 
    full_width = F, 
    position = "center"
  )
```


```{r w/o spline more covariates, out.width = '50%', fig.align = 'center', fig.show = 'hide', include = FALSE}
# Not include mutation; 
# include timing, surgery, gadx, cancer type and neoadjuvant
# stage and surgery and timing and gadx_norm (2, 5, 6)

# try all possible combinations
fixed_terms <- paste0(
  c(
    "I_HER2", "gadx_norm", "I_TNBC", "I_stage_34",
    "I_surgery", "I_neoadjuvant"
  ),
  collapse = " + "
)
optional_terms <- c(
  "yeardx_centered",
  # neoadjuvant, stage and surgery are correlated
  "I_neoadjuvant : I_stage_34", "I_surgery : I_stage_34",
  "I_surgery : I_neoadjuvant",
  # gadx related
  paste0(
    "gadx_norm : ",
    c(
     "I_TNBC", "I_stage_34",
      "I_surgery", "I_neoadjuvant"
    )
  ),
  # TNBC related
  paste0(
    "I_TNBC : ",
    c("I_stage_34", "I_surgery", "I_neoadjuvant")
  )
)
# all possible combination of interaction terms
optional_term_combn <- lapply(
  1:length(optional_terms), function(i) {
    combn(optional_terms, i,
      function(x) paste0(x, collapse = " + "),
      simplify = TRUE
    )
  }
) %>% 
  do.call(c, .) # combine list elements
# concatenate together
formula_taxol <- paste(
  paste("taxolpreg", fixed_terms, sep = " ~ "), 
  optional_term_combn,
  sep = " + "
)

M_taxol <- c(99, 99)
I_taxol <- c(NULL, NULL)
for (i in 1:length(formula_taxol)) {
  # fit model and extract std diff before and after
  .std_diff_tbl <- psw(
    data = as.data.frame(ps_data),
    form.ps = formula_taxol[i],
    weight = 'MW',
    std.diff = TRUE,
    V.name = tbl1_vars
  ) %>%
    extractStdDiff(.)
  # extract std diff after
  .std_diff_after_w_yeardx <- .std_diff_tbl$After
  .std_diff_after_wo_yeardx <- .std_diff_tbl$After[
    .std_diff_tbl$Vars != "yeardx_centered"]

  if (max(abs(.std_diff_after_w_yeardx)) < M_taxol[1]) {
    M_taxol[1] <- max(abs(.std_diff_after_w_yeardx))
    I_taxol[1] <- i
  }

  if (max(abs(.std_diff_after_wo_yeardx)) < M_taxol[2]) {
    M_taxol[2] <- max(abs(.std_diff_after_wo_yeardx))
    I_taxol[2] <- i
  }
}

# Best results without yeardx
cat(
  "When including year of diagnosis for the comparison of standardized difference before and after propensity score matching, the best formula is \n\n", formula_taxol[I_taxol[1]], "\n\n",
  "The largest standard difference is \n", M_taxol[1], "\n\n",  
  "When not including year of diagnosis for the comparison of standardized difference before and after propensity score matching, the best formula is \n\n", formula_taxol[I_taxol[2]], "\n\n",
  "The largest standard difference is \n", M_taxol[2], "\n"
)

ps_data %>% 
  group_by(year_grp, taxolpreg) %>% 
  summarise(n = n()) %>% 
  pivot_wider(names_from = taxolpreg, values_from = n)

```

I then tried to include more covariates and interaction terms in the model. All possible combinations of interaction terms were tried. When using ATE weight, the best PS matching model is: 

`r forms_taxol[227]`

<!-- The standardized difference table is: -->

```{r taxol ATE std diff table, eval = FALSE}
best_model_taxol_ate <-  psw(
  data = ps_data,
  form.ps = forms_taxol[227],
  weight = "ATE"
)
ps_data[, w_taxol_ate := best_model_taxol_ate$W]

std_diff_ATE_taxol[[227]] %>% 
  knitr::kable(align = "l", digits = 2, booktabs = TRUE,
               caption = "Standardized difference when using ATE weight") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F, position = "center")
```

When using MW weight, the best PS matching model is: 

`r forms_taxol[72]`. 

The standardized difference table is:

```{r taxol MW std diff table}
best_model_taxol_mw <- psw(
  data = ps_data,
  form.ps = forms_taxol[72],
  weight = "MW"
)
ps_data[, w_taxol_mw := best_model_taxol_mw$W]

std_diff_MW_taxol[[72]] %>% 
  knitr::kable(align = "l", digits = 2, booktabs = TRUE,
               caption = "Standardized difference when using MW weight") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F, position = "center")

```

## Propensity score model for GCSF

Among patients who received chemotherapy and gave single birth, only one patient didn't receive AC chemotherapy. The same covariates as above were fitted into this model.

```{r GCSF and AC chemotherapy}
ps_data[, .N, by = c("gfpreg", "acpreg")] %>% 
  knitr::kable(align = "l", digits = 2, booktabs = TRUE,
               col.names = c("GCSF", "AC chemotherapy", "N")) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F, position = "center")
```

```{r GCSF PS model, out.width = '50%', fig.align = 'center', fig.show = 'hide', include = FALSE}
# concatenate together
formula_gcsf <- paste(
  paste("gfpreg", fixed_terms, sep = " ~ "), 
  optional_term_combn,
  sep = " + "
)

M_gcsf <- c(99, 99)
I_gcsf <- c(NULL, NULL)

for (i in 1:length(formula_gcsf)) {
  # fit model and extract std diff before and after
  .std_diff_tbl <- psw(
    data = as.data.frame(ps_data),
    form.ps = formula_gcsf[i],
    weight = 'MW',
    std.diff = TRUE,
    V.name = tbl1_vars
  ) %>%
    extractStdDiff(.)
  # extract std diff after
  .std_diff_after_w_yeardx <- .std_diff_tbl$After
  .std_diff_after_wo_yeardx <- .std_diff_tbl$After[
    .std_diff_tbl$Vars != "yeardx_centered"]

  if (max(abs(.std_diff_after_w_yeardx)) < M_gcsf[1]) {
    M_gcsf[1] <- max(abs(.std_diff_after_w_yeardx))
    I_gcsf[1] <- i
  }

  if (max(abs(.std_diff_after_wo_yeardx)) < M_gcsf[2]) {
    M_gcsf[2] <- max(abs(.std_diff_after_wo_yeardx))
    I_gcsf[2] <- i
  }
}

cat(
  "When including year of diagnosis for the comparison of standardized difference before and after propensity score matching, the best formula is \n\n", formula_gcsf[I_gcsf[1]], "\n\n",
  "The largest standard difference is \n", M_gcsf[1], "\n\n",  
  "When not including year of diagnosis for the comparison of standardized difference before and after propensity score matching, the best formula is \n\n", formula_gcsf[I_gcsf[2]], "\n\n",
  "The largest standard difference is \n", M_gcsf[2], "\n"
)


```

When using ATE weight, the best PS matching model is: 

`r forms_gcsf[793]`. 

<!-- The standardized difference table is: -->

```{r GCSF ATE std diff table, eval = FALSE}
best_model_gcsf_ate <- psw(
  data = ps_data,
  form.ps = forms_gcsf[793],
  weight = "ATE"
)
ps_data[, w_gcsf_mw := best_model_gcsf_ate$W]

std_diff_ATE_gcsf[[793]] %>% 
  knitr::kable(align = "l", digits = 2, booktabs = TRUE,
               caption = "Standardized difference when using ATE weight") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F, position = "center")
```

When using MW weight, the best PS matching model is: 

`r forms_gcsf[580]`. 

The standardized difference table is:

```{r GCSF MW std diff table}
best_model_gcsf_mw <- psw(
  data = ps_data,
  form.ps = forms_gcsf[580],
  weight = "MW"
)
ps_data[, w_gcsf_mw := best_model_gcsf_mw$W]


std_diff_MW_gcsf[[580]] %>% 
  knitr::kable(align = "l", digits = 2, booktabs = TRUE,
               caption = "Standardized difference when using MW weight") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F, position = "center")

```

# Average treatment effect

Due to a low event rate. Statistical analysis will only be performed for the composite outcomes. There are two ways to define composite outcomes:

* Primary analysis: Composite outcome = missing only if all components are missing
* Sensitivity analysis: Composite outcome = missing if any of the components are missing


```{r composite outcomes missing}
rbind(
  factorStats(ps_data[, .(obstetrical)], useNA = "ifany"),
  factorStats(ps_data[, .(gestational)], useNA = "ifany"),
  factorStats(ps_data[, .(obstetrical_sens)], useNA = "ifany"),
  factorStats(ps_data[, .(gestational_sens)], useNA = "ifany")
) %>%
  knitr::kable(align = "l", digits = 2, booktabs = TRUE,
               caption = "Summary statistics for compostie outcomes") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F, position = "center")

```

## Average treatment effect of taxol on completed cases

```{r ATE taxol complete, fig.show = 'hide'}
# psw does not support interaction terms when calculating ATE
# based on forms_taxol[227]: taxolpreg ~ I_HER2 + gadx_norm + I_TNBC + I_stage_34 + I_surgery + I_neoadjuvant + I_neoadjuvant : I_stage_34 + gadx_norm : I_surgery + gadx_norm : I_stage_34 + I_TNBC : I_surgery
ps_data[, neo_stage := I_neoadjuvant * I_stage_34]
ps_data[, gadx_surg := gadx_norm * I_surgery]
ps_data[, gadx_stage := gadx_norm * I_stage_34]
ps_data[, tnbc_surg := I_TNBC * I_surgery]
ps_data[, surg_stage := I_surgery * I_stage_34]
ps_data[, gadx_tnbc := gadx_norm * I_TNBC]
ps_data[, tnbc_stage := I_TNBC * I_stage_34]
ps_data[, tnbc_neo := I_TNBC * I_neoadjuvant]
ps_data[, surg_neo := I_surgery * I_neoadjuvant]

# save(ps_data, file = file.path("RData", "ps_data.RData"))

taxol_best_formulas <- list(
  ATE = "taxolpreg ~ I_HER2 + gadx_norm + I_TNBC + I_stage_34 + I_surgery + I_neoadjuvant + 
                neo_stage + gadx_surg + gadx_stage + tnbc_surg",
  MW = "taxolpreg ~ I_HER2 + gadx_norm + I_TNBC + I_stage_34 + I_surgery + I_neoadjuvant + 
                neo_stage + gadx_surg + gadx_stage"
)

gcsf_best_formulas <- list(
  ATE = "gfpreg ~ I_HER2 + gadx_norm + I_TNBC + I_stage_34 + I_surgery + I_neoadjuvant + 
                surg_stage + surg_neo + gadx_stage + tnbc_surg + tnbc_stage + tnbc_neo",
  MW = "gfpreg ~ I_HER2 + gadx_norm + I_TNBC + I_stage_34 + I_surgery + I_neoadjuvant + 
                surg_stage + gadx_tnbc + tnbc_surg + tnbc_stage + tnbc_neo"
)

ATE_taxol_complete <- sapply(c("ATE", "MW"), function(w) {
  sapply(
    c("obstetrical", "gestational", "obstetrical_sens", "gestational_sens"), 
    function(outc) {
      .psw <- psw(data = as.data.frame(ps_data[!is.na(get(outc)), ]),
                  form.ps = taxol_best_formulas[[w]],
                  weight = w, 
                  wt = TRUE,
                  out.var = outc,
                  family = "binomial")
      .ATE <- extractATE(.psw) # return a data frame
      colnames(.ATE) <- c("est_risk", "std_risk", "est_rr", "std_rr", "est_or", "std_or")
      .ATE
      },
    simplify = FALSE, USE.NAMES = TRUE
  ) %>%
    rbindlist(., idcol = "outcome")
},
simplify = FALSE, USE.NAMES = TRUE)


# ATE_taxol_complete$ATE %>% 
#   knitr::kable(align = "l", digits = 2, booktabs = TRUE,
#                caption = "Average treatment effect of taxol on completed cases with ATE weight") %>%
#   kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F, position = "center")

ATE_taxol_complete$MW %>% 
  knitr::kable(align = "l", digits = 2, booktabs = TRUE,
               caption = "Average treatment effect of taxol on completed cases with MW weight") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F, position = "center")
```

## Average treatment effect of taxol on multiple imputed cases

```{r ATE taxol impute, results = 'hide'}
# ATE_taxol_imputed <- sapply(
#   c("ATE", "MW"), 
#   function(w) {
#     .data <- ps_data
#     sapply(
#       c("gestational", "obstetrical_sens", "gestational_sens"),
#       function(outc) {
#         .sub_data <- ps_data[, .SD, .SDcols = c(tbl1_vars, outc)]
#         .fit <- mice(
#           data = .sub_data,
#           m = 10, # number of MIs
#           maxit = 50, # iteration
#           seed = 42,
#           method = "pmm"
#         )
#         sapply(1:10, function(i) {
#             imputed_outc <- mice::complete(.fit, i)[[outc]]
#             .data[[outc]] <- imputed_outc
#             .psw <- psw(
#               data = as.data.frame(.data),
#               form.ps = taxol_best_formulas[[w]],
#               weight = w,
#               wt = TRUE, 
#               out.var = outc,
#               family = "binomial"
#             )
#             .ATE <- extractATE(.psw) # return a data frame
#             colnames(.ATE) <- c("est_risk", "std_risk", "est_rr", "std_rr", "est_or", "std_or")
#             .ATE
#           },
#           simplify = FALSE, USE.NAMES = TRUE) %>% 
#           rbindlist(., idcol = "i")
#       },
#       simplify = FALSE, USE.NAMES = TRUE) %>% 
#       rbindlist(., idcol = "outcome")
#   },
#   simplify = FALSE, USE.NAMES = TRUE
# )
# 
# save(ATE_taxol_imputed, file = "ATE_taxol_imputed.RData")

load(file.path("RData", "ATE_taxol_imputed.RData"))


```

```{r ATE taxol impute pool}
# pool MI results
# https://stats.idre.ucla.edu/wp-content/uploads/2016/02/multipleimputation.pdf
PoolEstimates <- function(Q, U) {
  # Q is point estimate
  # U is variance estimate
  if (length(Q) != length(U)) {
    stop("Q and U are of different length")
  }

  m <- length(Q)
  Q_bar <- mean(Q)
  U_bar <- mean(U) # within-imputation variance
  B <- var(Q) # between-imputation variance
  tt <- U_bar + (1 + 1 / m) * B # total variance
  # vm <- (m - 1) * (1 + U_bar / ((1 + 1 / m) * B)) ^ 2 # degree of freedom
  .tmp <- c(Q_bar, tt)
  names(.tmp) <- c("Q_bar", "total_variance")
  return(.tmp)
}

ATE_taxol_pool <- lapply(ATE_taxol_imputed, function(d) {
  d %>% group_by(outcome) %>% 
    summarise(est.risk = PoolEstimates(est_risk, std_risk)[1],
            std.risk = sqrt(PoolEstimates(est_risk, std_risk)[2]),
            est.rr = PoolEstimates(est_rr, std_rr)[1],
            std.rr = sqrt(PoolEstimates(est_rr, std_rr)[2]),
            est.or = PoolEstimates(est_or, std_or)[1],
            std.or = sqrt(PoolEstimates(est_or, std_or)[2]))
})

# ATE_taxol_pool$ATE[c(1, 3, 2), ] %>%
#   knitr::kable(align = "l", digits = 2, booktabs = TRUE,
#                col.names = c("outcome", "est_risk", "std_risk", "est_rr", "std_rr", "est_or", "std_or"),
#                caption = "Average treatment effect of taxol on imputed cases with ATE weight") %>%
#   kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F, position = "center")

ATE_taxol_pool$MW[c(1, 3, 2), ] %>%
  knitr::kable(align = "l", digits = 2, booktabs = TRUE,
               col.names = c("outcome", "est_risk", "std_risk", "est_rr", "std_rr", "est_or", "std_or"),
               caption = "Average treatment effect of taxol on imputed cases with MW weight") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F, position = "center")

```

## Average treatment effect of GCSF on completed cases

```{r ATE GCSF complete, fig.show = 'hide'}
ATE_gcsf_complete <- sapply(c("ATE", "MW"), function(w) {
  sapply(
    c("obstetrical", "gestational", "obstetrical_sens", "gestational_sens"), 
    function(outc) {
      .psw <- psw(data = as.data.frame(ps_data[!is.na(get(outc)), ]),
                  form.ps = gcsf_best_formulas[[w]],
                  weight = w, 
                  wt = TRUE,
                  out.var = outc,
                  family = "binomial")
      .ATE <- extractATE(.psw) # return a data frame
      colnames(.ATE) <- c("est_risk", "std_risk", "est_rr", "std_rr", "est_or", "std_or")
      .ATE
      },
    simplify = FALSE, USE.NAMES = TRUE
  ) %>%
    rbindlist(., idcol = "outcome")
},
simplify = FALSE, USE.NAMES = TRUE)

# ATE_gcsf_complete$ATE %>% 
#   knitr::kable(align = "l", digits = 2, booktabs = TRUE,
#                caption = "Average treatment effect of GCSF on completed cases with ATE weight") %>%
#   kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F, position = "center")

ATE_gcsf_complete$MW %>% 
  knitr::kable(align = "l", digits = 2, booktabs = TRUE,
               caption = "Average treatment effect of GCSF on completed cases with MW weight") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F, position = "center")

```

## Average treatment effect of GCSF on multiple imputed cases

```{r ATE GCSF impute, fig.show = 'hide'}
# ATE_gcsf_imputed <- sapply(
#   c("ATE", "MW"),
#   function(w) {
#     .data <- ps_data
#     sapply(
#       c("gestational", "obstetrical_sens", "gestational_sens"),
#       function(outc) {
#         .sub_data <- ps_data[, .SD, .SDcols = c(tbl1_vars, outc)]
#         .fit <- mice(
#           data = .sub_data,
#           m = 10, # number of MIs
#           maxit = 50, # iteration
#           seed = 42,
#           method = "pmm"
#         )
#         sapply(1:10, function(i) {
#             imputed_outc <- mice::complete(.fit, i)[[outc]]
#             .data[[outc]] <- imputed_outc
#             .psw <- psw(
#               data = as.data.frame(.data),
#               form.ps = gcsf_best_formulas[[w]],
#               weight = w,
#               wt = TRUE,
#               out.var = outc,
#               family = "binomial"
#             )
#             .ATE <- extractATE(.psw) # return a data frame
#             colnames(.ATE) <- c("est_risk", "std_risk", "est_rr", "std_rr", "est_or", "std_or")
#             .ATE
#           },
#           simplify = FALSE, USE.NAMES = TRUE) %>%
#           rbindlist(., idcol = "i")
#       },
#       simplify = FALSE, USE.NAMES = TRUE) %>%
#       rbindlist(., idcol = "outcome")
#   },
#   simplify = FALSE, USE.NAMES = TRUE
# )
# 
# save(ATE_gcsf_imputed, file = "ATE_gcsf_imputed.RData")

load(file.path("RData", "ATE_gcsf_imputed.RData"))


```

```{r}
ATE_gcsf_pool <- lapply(ATE_gcsf_imputed, function(d) {
  d %>% group_by(outcome) %>% 
    summarise(est.risk = PoolEstimates(est_risk, std_risk)[1],
            std.risk = sqrt(PoolEstimates(est_risk, std_risk)[2]),
            est.rr = PoolEstimates(est_rr, std_rr)[1],
            std.rr = sqrt(PoolEstimates(est_rr, std_rr)[2]),
            est.or = PoolEstimates(est_or, std_or)[1],
            std.or = sqrt(PoolEstimates(est_or, std_or)[2]))
})


# ATE_gcsf_pool$ATE[c(1, 3, 2), ] %>%
#   knitr::kable(align = "l", digits = 2, booktabs = TRUE,
#                col.names = c("outcome", "est_risk", "std_risk", "est_rr", "std_rr", "est_or", "std_or"),
#                caption = "Average treatment effect of GCSF on imputed cases with ATE weight") %>%
#   kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F, position = "center")


ATE_gcsf_pool$MW[c(1, 3, 2), ] %>%
  knitr::kable(align = "l", digits = 2, booktabs = TRUE,
               col.names = c("outcome", "est_risk", "std_risk", "est_rr", "std_rr", "est_or", "std_or"),
               caption = "Average treatment effect of GCSF on imputed cases with MW weight") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F, position = "center")

```

```{r, eval = FALSE}
# tbl3_outc <- c("sptb", "pprom", "gdm", "pih", "chorio", "sga", "apgar5v7", "anomalies", "gestational", "obstetrical")
tbl3_outc <- c("gestational", "obstetrical")

# outcome ~ taxol + 1 covariate
conditional_model_taxol <- sapply(tbl3_outc, function(outc) {
  w <- ps_data[["w_taxol_mw"]]
  sapply(tbl1_vars, function(x) {
    f <- as.formula(paste0(tbl3_outc, " ~ taxolpreg + ", x))
    .glm <- glm(f, family = binomial, data = ps_data, weights = w)
    return(summary(.glm))
  }, 
  simplify = FALSE, USE.NAMES = TRUE)
}, 
simplify = FALSE, USE.NAMES = TRUE)
# outcome ~ gcsf + 1 covariate
conditional_model_gcsf <- sapply(tbl3_outc, function(outc) {
  w <- ps_data[["w_gcsf_mw"]]
  sapply(tbl1_vars, function(x) {
    f <- as.formula(paste0(tbl3_outc, " ~ gfpreg + ", x))
    .glm <- glm(f, family = binomial, data = ps_data, weights = w)
    return(summary(.glm))
  }, 
  simplify = FALSE, USE.NAMES = TRUE)
}, 
simplify = FALSE, USE.NAMES = TRUE)

# calculate odds ratio of treatment effect
sapply(conditional_model_taxol, function(outc) {
  sapply(outc, function(.cov) {
    .coef <- .cov$coefficients["taxolpreg", c("Estimate", "Pr(>|z|)")]
    .coef[1] <- exp(.coef[1]) # odds ratio
    return(data.frame(odds.ratio = .coef[1],
                      p.value = .coef[2]))
  }, 
  simplify = FALSE, USE.NAMES = TRUE) %>% 
    rbindlist(., idcol = "Covariate")
},
simplify = FALSE, USE.NAMES = TRUE) %>% 
  rbindlist(., idcol = "Outcome") %>% 
  knitr::kable(align = "l", digits = 2, booktabs = TRUE,
               caption = "Odds ratio and p-value for taxol treatment",
               col.names = c("Outcome", "Covariate", "Odds ratio (Taxol)", "P value")) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F, position = "center") %>% 
  collapse_rows(columns = 1, valign = "top")

sapply(conditional_model_gcsf, function(outc) {
  sapply(outc, function(.cov) {
    .coef <- .cov$coefficients["gfpreg", c("Estimate", "Pr(>|z|)")]
    .coef[1] <- exp(.coef[1]) # odds ratio
    return(data.frame(odds.ratio = .coef[1],
                      p.value = .coef[2]))
  }, 
  simplify = FALSE, USE.NAMES = TRUE) %>% 
    rbindlist(., idcol = "Covariate")
},
simplify = FALSE, USE.NAMES = TRUE) %>% 
  rbindlist(., idcol = "Outcome") %>% 
  knitr::kable(align = "l", digits = 2, booktabs = TRUE,
               caption = "Odds ratio and p-value for gcsf treatment",
               col.names = c("Outcome", "Covariate", "Odds ratio (GCSF)", "P value")) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F, position = "center") %>% 
  collapse_rows(columns = 1, valign = "top")

```

```{r unused, out.width = '50%', fig.align = 'center', eval = FALSE, include = FALSE}
# w/ spline few covariates
# original scale
gadx_summary <- summary(ps_data$gadx)
spline_gadx <- bs(ps_data$gadx, knots = gadx_summary[c(2, 3, 5)]) # 1st quantile, median and 3rd quantile
colnames(spline_gadx) <- paste0("gadx_bs", 1:6)
ps_data <- cbind(ps_data, spline_gadx)
# normalized
gadx_norm_summary <- summary(ps_data$gadx_norm)
spline_gadx_norm <- bs(ps_data$gadx_norm, knots = gadx_norm_summary[c(2, 3, 5)])
colnames(spline_gadx_norm) <- paste0("gadx_norm_bs", 1:6)
ps_data <- cbind(ps_data, spline_gadx_norm)

psw_spline <- psw(
  data = as.data.frame(ps_data),
  form.ps = taxolpreg ~ stage_3_4 + gadx_norm_bs1 + gadx_norm_bs2 + gadx_norm_bs3 + gadx_norm_bs4 + gadx_norm_bs5 + gadx_norm_bs6,
  weight = 'ATE',
  std.diff = TRUE
)

std_diff_spline <- data.frame(
  Vars = row.names(psw_spline$std.diff.before),
  Before = psw_spline$std.diff.before[, 5],
  After = psw_spline$std.diff.after[, 5]
)
rownames(std_diff_spline) <- NULL

knitr::kable(std_diff_spline, align = "l", digits = 2, booktabs = TRUE,
             caption = "Standardized difference when using splined gestational age") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F, position = "center")

```

```{r unused code, eval = FALSE}
MI_outcomes <- c("sptb", "gdm", "pih", "chorio") # PPROM has too many missing; these are obstetrical outcomes
for (i in MI_outcomes) { # all other variables have no missing
  md.pattern(ps_data[, .SD, .SDcols = c(tbl_1_factor_vars, tbl_1_num_vars, "taxolpreg", i)])
}

# as.factor and normalization
ps_data[, (tbl_1_factor_vars) := lapply(.SD, as.factor), .SDcols = tbl_1_factor_vars]
ps_data[, (tbl_1_dummy_vars) := lapply(.SD, as.factor), .SDcols = tbl_1_dummy_vars]
# Formulas
formula_1 <- formula(paste0("taxolpreg ~ ", paste(tbl_1_factor_vars, collapse = " + "), " + ", paste(tbl_1_num_vars, collapse = " + ")))
formula_AIC <- stepAIC(glm(formula_dummy, family = binomial, data = ps_data), trace = FALSE)$formula
# taxolpreg ~ stage4 + cancer_HER2 + timing_neoadjuvant + gcsf_daily + gadx
# taxolpreg ~ stage4 + timing_neoadjuvant + surgpg0 + gadx
formula_AIC_plus <- formula("taxolpreg ~ stage1 + stage3 + stage4 + timing_neoadjuvant + surgpg0 + gadx")

# Fitted models
glm_model_full <- glm(formula_dummy, family = binomial, data = ps_data) # with all dummy variables
glm_model_AIC <- glm(formula_AIC, family = binomial, data = ps_data) # with only factor levels selected by AIC
glm_model_AIC_plus <- glm(formula_AIC_plus, family = binomial, data = ps_data) # with factors whose levels are selected by AIC
# CalculATE propensity score
ps_data[, ps_score_full := predict(glm_model_full, type = "response")]
ps_data[, ps_score_AIC := predict(glm_model_AIC, type = "response")]
ps_data[, ps_score_AIC_plus := predict(glm_model_AIC_plus, type = "response")]

calibration.plot(ps_data$taxolpreg, ps_data$ps_score_full) # ps_score_full performs the best
calibration.plot(ps_data$taxolpreg, ps_data$ps_score_AIC)
calibration.plot(ps_data$taxolpreg, ps_data$ps_score_AIC_plus)

prob.hist(ps_data$taxolpreg, ps_data$ps_score_full) # the same as above
prob.hist(ps_data$taxolpreg, ps_data$ps_score_AIC)
prob.hist(ps_data$taxolpreg, ps_data$ps_score_AIC_plus)


# taxolpreg ~ stage1 + stage3 + stage4 + cancer_HER2 + cancer_TNBC + timing_neoadjuvant + gcsf_daily + gcsf_depot + gadx
ps_data[, ps_score_AIC_plus := predict(glm_model_AIC_plus, type = "response")]
ps_data[, ps_score_AIC := predict(glm_model_AIC, type = "response")]
ps_data[, sptb1 := ifelse(is.na(sptb), 1, sptb)]
ps_data[, sptb0 := ifelse(is.na(sptb), 0, sptb)]
ps_data[, pprom1 := ifelse(is.na(pprom), 1, pprom)]
ps_data[, pprom0 := ifelse(is.na(pprom), 0, pprom)]
ps_data[, gdm1 := ifelse(is.na(gdm), 1, gdm)]
ps_data[, gdm0 := ifelse(is.na(gdm), 0, gdm)]
ps_data[, pih1 := ifelse(is.na(pih), 1, pih)]
ps_data[, pih0 := ifelse(is.na(pih), 0, pih)]
ps_data[, sga1 := ifelse(is.na(sga), 1, sga)]
ps_data[, sga0 := ifelse(is.na(sga), 0, sga)]
ps_data[, apgar1 := ifelse(is.na(apgar5v7), 1, apgar5v7)]
ps_data[, apgar0 := ifelse(is.na(apgar5v7), 0, apgar5v7)]
ps_data[, anomalies1 := ifelse(is.na(anomalies), 1, anomalies)]
ps_data[, anomalies0 := ifelse(is.na(anomalies), 0, anomalies)]

single_outcome <- c("sptb1", "sptb0", "pprom1", "pprom0", "gdm1", "gdm0", "pih1", "pih0", "sga1", "sga0", "apgar1",
                    "apgar0", "anomalies1", "anomalies0")
sensitivity_analysis <- setNames(data.frame(matrix(ncol = 2, nrow = 0)), c("Variable", "ATE"))
for (i in single_outcome) {
  sensitivity_analysis <- rbind(
    sensitivity_analysis,
    data.frame(Variable = i,
               ATE = weighted.mean(ps_data[[i]][ps_data$taxolpreg == 1], ps_data[taxolpreg == 1, 1 / ps_score_AIC_plus]) -
                 weighted.mean(ps_data[[i]][ps_data$taxolpreg == 0], ps_data[taxolpreg == 0, 1 / (1 - ps_score_AIC_plus)])))
}

sptb_p0 <- ggplot(ps_data, aes(x = factor(sptb0), y = ps_score_AIC_plus)) + geom_boxplot() + theme_bw() + labs(title = "", x = "NA as 0", y = "Propensity score")
sptb_p1 <- ggplot(ps_data, aes(x = factor(sptb1), y = ps_score_AIC_plus)) + geom_boxplot() + theme_bw() + labs(x = "NA as 1", y = "Propensity score")
grid.arrange(sptb_p0, sptb_p1, ncol = 2, nrow = 1)

pprom_p0 <- ggplot(ps_data, aes(x = factor(pprom0), y = ps_score_AIC_plus)) + geom_boxplot() + theme_bw() + labs(x = "NA as 0", y = "Propensity score")
pprom_p1 <- ggplot(ps_data, aes(x = factor(pprom1), y = ps_score_AIC_plus)) + geom_boxplot() + theme_bw() + labs(x = "NA as 1", y = "Propensity score")
grid.arrange(pprom_p0, pprom_p1, ncol = 2, nrow = 1)

gdm_p0 <- ggplot(ps_data, aes(x = factor(gdm0), y = ps_score_AIC_plus)) + geom_boxplot() + theme_bw() + labs(x = "NA as 0", y = "Propensity score")
gdm_p1 <- ggplot(ps_data, aes(x = factor(gdm1), y = ps_score_AIC_plus)) + geom_boxplot() + theme_bw() + labs(x = "NA as 1", y = "Propensity score")
grid.arrange(gdm_p0, gdm_p1, ncol = 2, nrow = 1)

pih_p0 <- ggplot(ps_data, aes(x = factor(pih0), y = ps_score_AIC_plus)) + geom_boxplot() + theme_bw() + labs(x = "NA as 0", y = "Propensity score")
pih_p1 <- ggplot(ps_data, aes(x = factor(pih1), y = ps_score_AIC_plus)) + geom_boxplot() + theme_bw() + labs(x = "NA as 1", y = "Propensity score")
grid.arrange(pih_p0, pih_p1, ncol = 2, nrow = 1)

sga_p0 <- ggplot(ps_data, aes(x = factor(sga0), y = ps_score_AIC_plus)) + geom_boxplot() + theme_bw() + labs(x = "NA as 0", y = "Propensity score")
sga_p1 <- ggplot(ps_data, aes(x = factor(sga1), y = ps_score_AIC_plus)) + geom_boxplot() + theme_bw() + labs(x = "NA as 1", y = "Propensity score")
grid.arrange(sga_p0, sga_p1, ncol = 2, nrow = 1)

apgar_p0 <- ggplot(ps_data, aes(x = factor(apgar0), y = ps_score_AIC_plus)) + geom_boxplot() + theme_bw() + labs(x = "NA as 0", y = "Propensity score")
apgar_p1 <- ggplot(ps_data, aes(x = factor(apgar1), y = ps_score_AIC_plus)) + geom_boxplot() + theme_bw() + labs(x = "NA as 1", y = "Propensity score")
grid.arrange(apgar_p0, apgar_p1, ncol = 2, nrow = 1)

anomalies_p0 <- ggplot(ps_data, aes(x = factor(anomalies0), y = ps_score_AIC_plus)) + geom_boxplot() + theme_bw() + labs(x = "NA as 0", y = "Propensity score")
anomalies_p1 <- ggplot(ps_data, aes(x = factor(anomalies1), y = ps_score_AIC_plus)) + geom_boxplot() + theme_bw() + labs(x = "NA as 1", y = "Propensity score")
grid.arrange(anomalies_p0, anomalies_p1, ncol = 2, nrow = 1)

# Check selected covariATEs
ggplot(ps_data, aes(x = factor(taxolpreg), fill = factor(stage))) + geom_bar(position = "fill") + theme_bw() +
  labs(title = "Compare distribution of cancer stages in different groups", x = "Taxol chemotherapy", y = "Percentage", fill = "Cancer stage")
ggplot(ps_data, aes(x = factor(taxolpreg), fill = factor(cancer_type))) + geom_bar(position = "fill") + theme_bw() +
  labs(title = "Compare distribution of cancer types in different groups", x = "Taxol chemotherapy", y = "Percentage", fill = "Cancer type")
ggplot(ps_data, aes(x = factor(taxolpreg), fill = factor(chemo_timing))) + geom_bar(position = "fill") + theme_bw() +
  labs(title = "Compare distribution of chemotherapy timing in different groups", x = "Taxol chemotherapy", y = "Percentage", fill = "Timing")
ggplot(ps_data, aes(x = factor(taxolpreg), fill = factor(gcsf))) + geom_bar(position = "fill") + theme_bw() +
  labs(title = "Compare distribution of GCSF treatment in different groups", x = "Taxol chemotherapy", y = "Percentage", fill = "GCSF")
ggplot(ps_data, aes(x = factor(taxolpreg), y = gadx)) + geom_boxplot() + theme_bw() +
  labs(title = "Compare distribution of gestational age in different groups", x = "Taxol chemotherapy", y= "Gestational age at diagnosis")
# ggplot(ps_data, aes(x = gadx, color = factor(taxolpreg))) + geom_density(aes(fill = factor(taxolpreg)), alpha = 0.5)
# Standardized difference
std_diff_table <- bind_rows(
  StdDiff(ps_data$taxolpreg, ps_data$stage, type = "cATEgory"),
  StdDiff(ps_data$taxolpreg, ps_data$cancer_type, type = "cATEgory"),
  StdDiff(ps_data$taxolpreg, ps_data$chemo_timing, type = "binary"),
  StdDiff(ps_data$taxolpreg, ps_data$gcsf, type = "cATEgory"),
  StdDiff(ps_data$taxolpreg, ps_data$gadx, type = "numeric")
)
std_diff_table <- bind_cols(Variable = c("stage", "cancer type", "chemotherapy timing", "GCSF", "gadx"),
                            std_diff_table)

# Weighted PS matching
ps_matching_AIC <- psw(
  as.data.frame(ps_data),
  form.ps = formula_AIC, std.diff = TRUE,
  # stage4 + cancer_HER2 + timing_neoadjuvant + gcsf_daily + gadx
  V.name = c("stage4", "cancer_HER2", "timing_neoadjuvant", "gcsf_daily", "gadx"),
  mirror.hist = TRUE, weight = "MW", family = "binomial"
) # psw function only contains the vars used and cATEgory vars need to be converted to dummy
ps_matching_AIC_plus <- psw(
  as.data.frame(ps_data), form.ps = formula_AIC_plus, weight = "MW", std.diff = TRUE,
  V.name = c("stage1", "stage3", "stage4", "cancer_HER2", "cancer_TNBC", "timing_neoadjuvant",
             "gcsf_daily", "gcsf_depot", "gadx"),
  mirror.hist = TRUE, family = "binomial" # add.weight = TRUE,
)
# unweighted PS score matching in three ways
unweighted_nn_1 <- matchit(ps_formula, data = ps_model_data, method = "nearest", ratio = 1)
unweighted_nn_2 <- matchit(ps_formula, data = ps_model_data, method = "nearest", ratio = 2)
unweighted_nn_3 <- matchit(ps_formula, data = ps_model_data, method = "nearest", ratio = 2, replace = TRUE)

```





























