---
title: "propensity score final"
output: 
  html_document:
    number_sections: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  include = TRUE, warning = FALSE, message = FALSE, echo = FALSE
)
```

```{r import data, include = FALSE}
packages_list <- c(
  "data.table", "tidyverse", "kableExtra", "PSW", "MASS", "sandwich", "mice"
) # 'MatchIt',
new_packages <- packages_list[
  !(packages_list %in% installed.packages()[, "Package"])
]
if (length(new_packages)) install.packages(new_packages)
suppressMessages(lapply(packages_list, library, character.only = TRUE))

select <- dplyr::select
rbind <- base::rbind
kable <- knitr::kable

extractATE <- function(.psw) {
  .wt_estimator <- .psw[
    c(
      "est.risk.wt", "std.risk.wt", "est.rr.wt", "std.rr.wt",
      "est.or.wt", "std.or.wt"
    )
  ] # list
  .wt_estimator <- data.frame(
    matrix(
      unlist(.wt_estimator),
      ncol = length(.wt_estimator)
    )
  )
  colnames(.wt_estimator) <- c(
    "est.risk.wt", "std.risk.wt", "est.rr.wt", "std.rr.wt",
    "est.or.wt", "std.or.wt"
  )
  .wt_estimator
}

extractStdDiff <- function(.psw) {
  .df <- data.frame(
    Vars = row.names(.psw$std.diff.before),
    Before = unname(.psw$std.diff.before[, 5]),
    After = unname(.psw$std.diff.after[, 5])
  )
  .df
}

# https://github.com/cran/PSW/blob/master/R/psw-internal.R
diff.plot <- function(diff.before, diff.after, name, weight, title = NULL) {
  # plot standardize difference for before and after weighting
  #
  # Args:
  #   diff.before: standardized mean differece before weighting
  #   diff.after: standardized mean difference after weighting
  #   name: covariate name
  #   weight: propensity score weighting method
  # Returns
  #   A plot is generated

  par(las = 1, lwd = 2, mar = c(5, max(nchar(name)), 4, 2), bty = "n")

  x.range <- range(c(diff.before, diff.after))
  y.range <- c(1, length(name))
  ord <- order(diff.before, decreasing = T)
  plot(
    x = x.range, y = y.range, xaxt = "n", yaxt = "n", type = "n",
    xlim = x.range, ylim = y.range,
    ylab = "", xlab = "Standardized mean difference",
    main = title
  )
  axis(side = 1, at = pretty(x.range))
  axis(side = 2, at = length(name):1, labels = name[ord], tick = F)
  abline(v = 0, lty = 1, col = "gray")
  abline(v = -10, lty = 3, lwd = 1)
  abline(v = 10, lty = 3, lwd = 1)
  points(y = length(name):1, x = diff.before[ord], pch = 4)
  points(y = length(name):1, x = diff.after[ord], pch = 21)
  legend("topleft", legend = c("Unadjusted", weight), pch = c(4, 21))
}

## MI estimates of treatment
# pool MI results
# https://stats.idre.ucla.edu/wp-content/uploads/2016/02/multipleimputation.pdf
PoolEstimates <- function(Q, U) {
  # Q is point estimate
  # U is variance estimate
  if (length(Q) != length(U)) {
    stop("Q and U are of different length")
  }
  m <- length(Q)
  Q_bar <- mean(Q)
  U_bar <- mean(U) # within-imputation variance
  B <- var(Q) # between-imputation variance
  tt <- U_bar + (1 + 1 / m) * B # total variance
  .tmp <- c(Q_bar, tt)
  names(.tmp) <- c("Q_bar", "total_variance")
  return(.tmp)
}

# load data
redcap_data <- fread(
  file.path(
    "C:/Users/sren/Dropbox (Partners HealthCare)/BOC shared/Chemo during pregnancy (Sella)/data/2020-10-23",
    "cleaned_data.csv"
  )
)
ps_data <- redcap_data[
  chemopg == 1 & pregnancy_outcome == "single" &
    !is.na(gadx),
]

best_formula_taxol <- "taxolpreg ~ I_HER2 + gadx_norm + I_TNBC + I_stage_34 +
I_surgery + I_neoadjuvant + neo_stage + gadx_surg + gadx_stage"

best_formula_gcsf <- "gfpreg ~ I_HER2 + gadx_norm + I_TNBC + I_stage_34 +
I_surgery + I_neoadjuvant + surg_stage + gadx_tnbc + tnbc_surg +
tnbc_stage + tnbc_neo"

tbl1_vars <- c(
  "age_norm", "gadx_norm", "I_stage_34",
  "I_TNBC", "I_HER2", "I_neoadjuvant", "I_surgery"
)
```

```{r modify table 1 vars for ps model}
# collapse factor levels, normalized numeric vars, and splined them
# normalized age
ps_data[, age_norm := (agedx - mean(agedx)) / sd(agedx)]
# normalized gestational age
ps_data[, gadx_norm := (gadx - mean(gadx)) / sd(gadx)]
# cancer stage
ps_data[, I_stage_34 := ifelse(stage %in% 3:4, 1, 0)] # collapse 1-2, and 3-4
# cancer type
ps_data[, I_TNBC := ifelse(cancer_type == "TNBC", 1, 0)]
ps_data[, I_HER2 := ifelse(cancer_type == "HER2", 1, 0)]
# BRCA1/2 
ps_data[, I_mutation := ifelse(mutation == "Y", 1, 0)] # collapse 0 and unknown
# chemotherapy timing
ps_data[, I_neoadjuvant := ifelse(chemo_timing == "neoadjuvant", 1, 0)]
# surgery
ps_data[, I_surgery := surgpg]

# psw does not support interaction terms when calculating ATE
# based on forms_taxol[227]: taxolpreg ~ I_HER2 + gadx_norm + I_TNBC + 
# I_stage_34 + I_surgery + I_neoadjuvant + I_neoadjuvant : I_stage_34 +
# gadx_norm : I_surgery + gadx_norm : I_stage_34 + I_TNBC : I_surgery
ps_data[, neo_stage := I_neoadjuvant * I_stage_34]
ps_data[, gadx_surg := gadx_norm * I_surgery]
ps_data[, gadx_stage := gadx_norm * I_stage_34]
ps_data[, tnbc_surg := I_TNBC * I_surgery]
ps_data[, surg_stage := I_surgery * I_stage_34]
ps_data[, gadx_tnbc := gadx_norm * I_TNBC]
ps_data[, tnbc_stage := I_TNBC * I_stage_34]
ps_data[, tnbc_neo := I_TNBC * I_neoadjuvant]
ps_data[, surg_neo := I_surgery * I_neoadjuvant]
```

In randomized experiments, subjects are randomly assigned to each treatment group so that each covariate is balanced across groups. However, in observational studies, the assignment of treatments is not random. Some of the covariates could affect the decision of treatment, which causes the observed difference in the treatment outcome can not be fully attributed to the treatment effect. Propensity score matching is a method that attempts to estimate the probability of receiving treatment for each subject. Based on the estimated treatment probability, subjects are assigned different weights to reduce this bias. The difference in the treatment outcome is then comparable.

# Propensity score matching model

In our study, there are two treatment effects that we want to evaluate: taxanes and colony stimulating factors (CSF). For each treatment, we built a logistic regression, with the treatment as the dependent variable and patient characteristics as confounders. By fitting this model, we obtained the probability of receiving taxane/CSF for each subject, and rebalanced the confounders accordingly. The confounders are selected from Table 1, which include: age at diagnosis, gestational age at diagnosis, cancer stage, cancer type, timing of chemotherapy and whether they had breast surgery during pregnancy. The performance of rebalancing were evaluated by the standardized difference of each confounder. Standardized difference is a measure of the effect size of covariates between two groups. Generally speaking, it calculates the difference in means divided by standard error. There is no specific threshold, but an absolute effect size between 0\% and 20\% is usually considered small. 

We also added several interaction terms to the logistic models. To decide which interaction terms should be added, we listed all the possible two-way interaction combinations, fitted corresponding models and compared model performance by the largest standardized difference in each model. There are also several choices of weighting shcemes. We compared two weights in this study: average treatment effect weight (ATE) and matching weight (MW). For both treatments, MW provides smaller overall standardized differences, so we use MW in all the following results. The formulas we used to get the best performances are:

For taxane: I(taxane = 1) ~ I(type = HER2) + I(type = TNBC) + gestational age + I(stage = 3,4) + I(surgery = 1) + I(timing = neoadjuvant) + I(timing = neoadjuvant) \* I(stage = 3,4) + gestational age \* I(surgery = 1) + gestational age \* I(stage = 3, 4)

For CSF: I(CSF = 1) ~ I(type = HER2) + I(type = TNBC) + gestational age + I(stage = 3,4) + I(surgery = 1) + I(timing = neoadjuvant) + I(surgery = 1) \* I(stage = 3,4) + gestational age \* I(type = TNBC) + I(type = TNBC) \* I(surgery = 1) + I(type = TNBC) \* I(stage = 3,4) + I(type = TNBC) \* I(timing = neoadjuvant)

```{r ps model, fig.show = "hide" } 
best_model_taxol <- psw(
  data = ps_data,
  form.ps = best_formula_taxol,
  weight = "MW",
  std.diff = TRUE,
  V.name = tbl1_vars
)

ps_data[, w_taxol_mw := best_model_taxol$W]

best_model_gcsf <- psw(
  data = ps_data,
  form.ps = best_formula_gcsf,
  weight = "MW",
  std.diff = TRUE,
  V.name = tbl1_vars
)

ps_data[, w_gcsf_mw := best_model_gcsf$W]

# output of glm model
taxol_coefs <- as.data.frame(summary(best_model_taxol$ps.model)$coefficients)
taxol_coefs$covariate <- c(
  "Intercept", "HER2", "Gestational age", "TNBC", "Stage", "Surgery",
  "Neoadjuvant", "Neoadjuvant*Stage", "Gestational age*Surgery",
  "Gestational age*Stage"
)
taxol_coefs %>%
  mutate(
    est.or = exp(Estimate),
    lower = exp(Estimate - 1.96 * `Std. Error`),
    upper = exp(Estimate + 1.96 * `Std. Error`),
  ) %>%
  select(covariate, est.or, lower, upper, `Pr(>|z|)`) %>%
  kable(
    digits = 2,
    col.names = c(
      "Covariate", "Estimated odds ratio", "Lower 95% CI",
      "Upper 95% CI", "P-value"
    ),
    caption = "Estimated coefficients for the taxane propensity score model"
  ) %>%
  kable_classic(
    full_width = F,
    position = "center",
    latex_options = "scale_down",
    font_size = 12
  )

gcsf_coefs <- as.data.frame(summary(best_model_gcsf$ps.model)$coefficients)
gcsf_coefs$covariate <- c(
  "Intercept", "HER2", "Gestational age", "TNBC", "Stage", "Surgery",
  "Neoadjuvant", "Surgery*Stage", "Gestational*TNBC", "TNBC*Surgery",
  "TNBC*Stage", "TNBC*Neoadjuvant"
)
gcsf_coefs %>%
  mutate(
    est.or = exp(Estimate),
    lower = exp(Estimate - 1.96 * `Std. Error`),
    upper = exp(Estimate + 1.96 * `Std. Error`),
  ) %>%
  select(covariate, est.or, lower, upper, `Pr(>|z|)`) %>%
  kable(
    digits = 2,
    col.names = c(
      "Covariate", "Estimated odds ratio", "Lower 95% CI",
      "Upper 95% CI", "P-value"
    ),
    caption = "Estimated coefficients for the CSF propensity score model"
  ) %>%
  kable_classic(
    full_width = F,
    position = "center",
    latex_options = "scale_down",
    font_size = 12
  )
```

```{r std diff plots}
diff.plot(
  diff.before = best_model_taxol$std.diff.before[, "std.diff.pct"],
  diff.after = best_model_taxol$std.diff.after[, "std.diff.pct"],
  name = c(
    "Age",
    "Gestational age",
    "Stage",
    "TNBC",
    "HER2",
    "Neoadjuvant",
    "Surgery"
  ),
  weight = "MW",
  title = "Adjusted for taxane"
)

diff.plot(
  diff.before = best_model_gcsf$std.diff.before[, "std.diff.pct"],
  diff.after = best_model_gcsf$std.diff.after[, "std.diff.pct"],
  name = c(
    "Age",
    "Gestational age",
    "Stage",
    "TNBC",
    "HER2",
    "Neoadjuvant",
    "Surgery"
  ),
  weight = "MW",
  title = "Adjusted for CSF"
)
```

# Average treatment effect model

The treatment effect of taxane\/CSF to obstetrical\/gestational outcomes could be evaluated by fitting separate weighted logistic regression models. In each model, we evaluate the effect of one treatment on one outcome. Each subject is assigned a weight based on its likelihood to receive treatment under investigation. Since there are many missingness in individual outcomes, we only evaluate treatment effect on composite obstetrical\/gestational outcomes. There are four ways to define composite outcomes:

1. If any individual outcome is positive, the composite outcome is positive; if any individual outcome is negative and no outcome is positive, the composite outcome is negative; else it's missing
2. If any individual outcome is positive, the composite outcome is positive; if all individual outcomes are negative, the composite outcome is negative; else it's missing
3. Generate composite outcomes the same way as method 1, then use multiple imputation (MI) method to impute missingness. For each composite outcome, 10 copies of the original dataset is created, then a regression model is fitted to predict missing values from complete cases. This process is iterated 25 times, updating regression parameters after each iteration. The formula is: \
composite outcome ~ age + gestational age + I(stage = 3,4) + I(type = TNBC) + I(type = HER2) + I(timing = neoadjuvant) + I(surgery)
4. Generate composite outcomes the same way as method 2, then use MI method to impute missingness

```{r create composite outcomes, include = FALSE}
## MI
mice_fit_ls <- sapply(
  c("obstetrical", "gestational", "obstetrical_sens", "gestational_sens"),
  function(outc) {
    vars <- c(tbl1_vars, outc)
    .sub_data <- ps_data[, .SD, .SDcols = vars]
    # mice object
    .fit <- mice(
      data = .sub_data,
      m = 10, # number of MIs
      maxit = 25, # iteration
      seed = 42,
      method = "pmm"
    )
    # imputed datasets
    imputed <- mice::complete(.fit, "all") # a list
    # add two treatments two the list
    imputed <- lapply(imputed, function(.data) {
      .data$taxolpreg <- ps_data$taxolpreg
      .data$gfpreg <- ps_data$gfpreg
      return(.data)
    })
    return(imputed)
  },
  simplify = FALSE,
  USE.NAMES = TRUE
)
```

```{r ATE model}
ATE_composite <- sapply(
  c("obstetrical", "gestational", "obstetrical_sens", "gestational_sens"),
  function(outc) {
    sapply(c("taxolpreg", "gfpreg"), function(trt) {
      w <<- if (trt == "taxolpreg") ps_data$w_taxol_mw else ps_data$w_gcsf_mw
      # outc ~ trt
      .glm <- glm(
        formula = paste0(outc, " ~ ", trt),
        family = binomial,
        data = ps_data,
        weights = w
      )
      # extract coefs
      .log_odds <- summary(.glm)$coefficients[trt, "Estimate"]
      .log_odds_var <- sandwich(.glm)[trt, trt]
      return(data.frame(
        est.log.or = .log_odds,
        std.log.or = sqrt(.log_odds_var)
      ))
    }, simplify = FALSE, USE.NAMES = TRUE) %>%
      rbindlist(., idcol = "Treatment")
  },
  simplify = FALSE, USE.NAMES = TRUE
) %>%
  rbindlist(., idcol = "Outcome") %>%
  mutate(
    est.or = exp(est.log.or),
    lower = exp(est.log.or - 1.96 * std.log.or),
    upper = exp(est.log.or + 1.96 * std.log.or)
  ) %>%
  arrange(Treatment, Outcome)

ATE_composite_MI <- sapply(
  c("obstetrical", "gestational", "obstetrical_sens", "gestational_sens"),
  function(outc) {
    .ls <- mice_fit_ls[[outc]]
    # tmp is a dataframe, with 10 iterations and 2 treatments
    tmp <- lapply(.ls, function(.data) {
      # taxane
      tmp1 <- {
        .glm <- glm(
          paste0(outc, " ~ taxolpreg"),
          family = binomial,
          data = .data,
          weights = ps_data$w_taxol_mw
        )
        .log_odds <- summary(.glm)$coefficients["taxolpreg", "Estimate"]
        .log_odds_var <- sandwich(.glm)["taxolpreg", "taxolpreg"]
        # return
        data.frame(
          Treatment = "taxolpreg",
          log.or = .log_odds,
          log.or.var = .log_odds_var
        )
      }
      # CSF
      tmp2 <- {
        .glm <- glm(
          paste0(outc, " ~ gfpreg"),
          family = binomial,
          data = .data,
          weights = ps_data$w_gcsf_mw
        )
        .log_odds <- summary(.glm)$coefficients["gfpreg", "Estimate"]
        .log_odds_var <- sandwich(.glm)["gfpreg", "gfpreg"]
        # return
        data.frame(
          Treatment = "gfpreg",
          log.or = .log_odds,
          log.or.var = .log_odds_var
        )
      }
      # combine taxane and CSF
      return(
        rbind(tmp1, tmp2)
      )
    }) %>%
      rbindlist(., idcol = "Iter")
    # combine stats over 10 imputed datasets
    tmp <- tmp %>%
      group_by(Treatment) %>%
      summarise(
        est.log.or = PoolEstimates(log.or, log.or.var)[1],
        std.log.or = sqrt(PoolEstimates(log.or, log.or.var)[2])
      ) %>%
      mutate(
        est.or = exp(est.log.or), # convert from coef to odds ratio
        lower = exp(est.log.or - 1.96 * std.log.or),
        upper = exp(est.log.or + 1.96 * std.log.or)
      ) %>%
      as.data.frame()
    return(tmp)
  },
  simplify = FALSE,
  USE.NAMES = TRUE
) %>%
  rbindlist(., idcol = "Outcome") %>%
  mutate(Outcome = paste0(Outcome, "_mi")) %>%
  arrange(Outcome, Treatment)

ATE_overall <- rbind(
  ATE_composite %>%
    select(Outcome, Treatment, est.or, lower, upper),
  ATE_composite_MI %>%
    select(Outcome, Treatment, est.or, lower, upper)
)

# For taxane
# taxane and obstetrical
ATE_overall %>%
  filter(Treatment == "taxolpreg" & str_detect(Outcome, "^obs")) %>%
  select(Outcome, est.or, lower, upper) %>%
  kable(
    digits = 2,
    col.names = c(
      "Outcome", "Estimated odds ratio", "Lower 95% CI", "Upper 95% CI"
    ),
    caption = "Estimated odds ratio for taxane chemotherapy on composite 
    obstetrical outcomes"
  ) %>%
  kable_classic(
    full_width = F,
    position = "center",
    latex_options = "scale_down",
    font_size = 12
  )
# taxane and gestational
ATE_overall %>%
  filter(Treatment == "taxolpreg" & str_detect(Outcome, "^ges")) %>%
  select(Outcome, est.or, lower, upper) %>%
  kable(
    digits = 2,
    col.names = c(
      "Outcome", "Estimated odds ratio", "Lower 95% CI", "Upper 95% CI"
    ),
    caption = "Estimated odds ratio for taxane chemotherapy on composite 
    gestational outcomes"
  ) %>%
  kable_classic(
    full_width = F,
    position = "center",
    latex_options = "scale_down",
    font_size = 12
  )
# GCSF and obstetrical
ATE_overall %>%
  filter(Treatment == "gfpreg" & str_detect(Outcome, "^obs")) %>%
  select(Outcome, est.or, lower, upper) %>%
  kable(
    digits = 2,
    col.names = c(
      "Outcome", "Estimated odds ratio", "Lower 95% CI", "Upper 95% CI"
    ),
    caption = "Estimated odds ratio for CSF chemotherapy on composite 
    obstetrical outcomes"
  ) %>%
  kable_classic(
    full_width = F,
    position = "center",
    latex_options = "scale_down",
    font_size = 12
  )
# GCSF and gestational
ATE_overall %>%
  filter(Treatment == "gfpreg" & str_detect(Outcome, "^ges")) %>%
  select(Outcome, est.or, lower, upper) %>%
  kable(
    digits = 2,
    col.names = c(
      "Outcome", "Estimated odds ratio", "Lower 95% CI", "Upper 95% CI"
    ),
    caption = "Estimated odds ratio for CSF chemotherapy on composite 
    gestational outcomes"
  ) %>%
  kable_classic(
    full_width = F,
    position = "center",
    latex_options = "scale_down",
    font_size = 12
  )
```




